Mitigating Interest Flooding DDoS Attacks in Named Data Networking

### 2012-12-05 04-design.tex
First part of the design section semi-done
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@855 df1a90cd-247b-4364-bdaf-287a704eb499

The presented methods rely on the core NDN principle of the flow balance between Interest and Data packets: Data packet can be forwarded only in response to the incoming Interest and each Interest is satisfied by at most one Data packet.
As a result of this principle, all traffic on NDN is ``receiver''-driven, where the ``receiver'' is either a user or any router on the path between the user and a Data producer.
Therefore, any node can limit the number of forwarded Interests, effectively limiting the amount of the retrieved Data.
At the same time, each forwarded Interest can be used to build up various data plane performance statistics, such as per-incoming interface ratios of satisfied Interests.

% Combining per-incoming interface data plane performance statistics with

% Disclaimer
% The presented solutions to mitigate Interest flooding DDoS attacks

% In particular, we explore different mechanisms to app

% Basic limits, probablity based on statistics, and dynamic window adjustment

@@

@@

@@

>>>>

%   - Naive approach to Interest Limits (that is called physical limits everywhere else)
%     Example how this can be implemented in a simple way
%     Baseline solution
%   - Interest limits with "fair" queuing
%     Improving problem of simple limits (no single face can dominate), but doesn't solve the proble

% - Attack mitigation
%   - Per-incoming interface Interest statistics

%   - Dynamic Interests limit adjustments

%   - Probabilistic Interest accept

Our methods to mitigate Interest flooding attack rely on the fundamental principle of the NDN architecture: the flow balance between Interest and Data packets: one Interest packet (the only communication initiator) can be satisfied with at most one Data packet.
Because NDN is host-to-host architecture (as opposed to end-to-end in the current IP Internet), the flow balance principle allow any entity on the network, end-hosts and routers, control what and how much Data they want to receive.
Therefore, any node can limit the number of forwarded Interests, effectively limiting the amount of the retrieved Data.
At the same time, each forwarded Interest can be used to build up various data plane performance statistics, such as per-incoming interface ratios of satisfied Interests.

\subsection{Na\"{i}ve attack mitigation}

% In a normal network operation, the size of Interests packets is supposed to be significantly smaller than the size of the requested Data.

% All communication in NDN network is host-to-host and receiver-driven.

The most obvious and na\"{i}ve method to mitigate the Interest flooding attack is to limit the number of forwarded Interests out of each interface. 
That is, if it is known that the amount of already forwarded Interests can fully utilize the downstream link, an NDN node---an end-user or an NDN router---has absolutely no point in forwarding new incoming Interests and creating corresponding PIT entries.
For example, if a router A on Fig.~\ref{fig:flooding example} already forwarded 125 Interests requesting 1000-byte Data packets, any new incoming Interests can be almost safely dropped, provided the link capacity between A--B is 100~Mbps and delay is 10~ms.
In other words, 125 Data packets returned by router B will fully utilize the link ($125 \times 1000 \mathrm{~bytes} \approx 10\mathrm{~ms} \times 100\mathrm{~Mbps}$) and any excess Data would be dropped.

\subsubsection{Interest limits (physical limits)}
\label{sec:physical limits}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Interest limits with per-interface fairness}
\label{sec:queuing}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Intelligent attack mitigation}

About additional knowledge about interests.

Another unique feature of the NDN architecture is stateful packet forwarding~\cite{adaptive-forwarding} and symmetric flow of Interests and Data packets.  Every time Interests is send out to the Interface, the router will be able to learn various parameters about this Interest.

From the DDoS mitigation point of view, one of the important parameters is to determine from which interface

In order to mitigate a denial of service attack it is vital to know the origin of the attack.  I the origin of the attack is known, then the attack can be easily suppressed and eventually localized.  In our definition of the attack is that

interest of the attackers are not getting satisfied. 
Every router can learn fro m which interfaces attack traffic is coming and do certain actions to limit forwarding of suspicious interest.   In other words, without trying to distinguish legitimate and malicious interests on the basis of the Interest prefix, routers distinguish them based on recent performance.

\subsubsection{Data plane performance tracking}
\label{sec:stats}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Probabilistic Interest accept}
\label{sec:probabilistic}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Dynamic adjustments of window-based limits}
\label{sec:dynamic limits}

### 2012-12-05 algos/physical-limits.tex
Adding fair-queueing algorithm description
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@815 df1a90cd-247b-4364-bdaf-287a704eb499
Correcting algorithms. Far far far from perfect...
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@849 df1a90cd-247b-4364-bdaf-287a704eb499
First part of the design section semi-done
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@855 df1a90cd-247b-4364-bdaf-287a704eb499

\subsection{Physical limits}

The requirement to send an Interest in order to receive Data packet, provides an NDN consumer a unique opportunity to request the right amount of Data.
Moreover, the same opportunity to control the amount of data flow is given not only to consumers, but all routers between consumer and producer (or nearby caches).  
In other words, every node, either a consumer or an intermediate router, is able to control how much data it wants to receive by limiting the number of forwarded Interests.

The limitation can implemented in a number of different ways, including leaky bucket scheduling and window-based flow control.
We decided to following TCP-like window-based flow control and applied the sliding window approach to implement Interest limits.

The size of the window defines how many Interests can be send out before Interests get satisfied or expired.
From the one hand, this size should be large enough to ``fill the pipe,'' meaning that a node needs to send enough Interests to receive Data at full capacity of the incoming link.
On the other hand, the window's size should not be too large to avoid excessive buffering and congestion of the Data packet.
Thus, the ideal size for such a window need to be defined proportional to link's bandwidth-delay product~\cite{tcp-survey}.
With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:

\[
\mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\]

Note that the value of \textit{Delay} is not known a priory and varies between different Interest-Data flows.
However, we do not need to know the exact value of the delay and can set it as an average round trip delay among all flows (with a reasonable filtering of outliers).
This way, the statistical traffic multiplexing with link-level buffering will allow full utilization of the downstream link.
Exactly the same reasoning can be applied to the \textit{Data packet size} parameter, which can also be set to an average observed Data packet size.

Unlike rate-based approaches, window-based limiting does not require precise knowledge about the rate, as well does not need precise scheduling mechanisms.
Like in TCP, the window-based flow is self-clocking, easily adjusting itself to any traffic patterns.

>>>>

% \subsection{Physical limits}
% \label{sec:physical limits}

One of the promising and simple ways to set and enforce limit on the number of forwarded Interests is a variation of the well-known bag of tokens algorithm, illustrated on Fig.~\ref{fig:bag of tokens}.
Each router implementing the algorithm maintains a ``bag'' with a limited number of tokens for every outgoing interface.
Whenever an Interest is ready to be forwarded, the router ``borrows'' a token from the outgoing interface's bag, dropping the Interest if the bag is empty.
At the same time, when the requested Data comes back or the Interest times out, the token is returned to the appropriate bag based on the information in PIT, ready to be borrowed again by a new incoming Interset.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.7]{token-bag}
  \caption{Bag of tokens algorithm: tokens borrowed when Interest is forwarded and returned when Data comes back or Interest times out}
  \label{fig:bag of tokens}
\end{figure}

Since the number of tokens in the bag defines how many Interests are forwarded and not yet satisfied or timed out (pending Interests), this number should be large enough to fully utilize the network and small enough to avoid excessive buffering and drops at the upstream routers.
The ideal value for the number of tokens (interest limit) for each individual link would be defined proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}.
This can be formalized in the following way:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:

\begin{equation}
\mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

In this equation, \emph{Delay} is the expected time of Interest being satisfied and \emph{Data packet size} is the size of the returned Data packet.
Although both these values are not known a priory and vary between different Interest-Data exchanges, it is not really necessary to know their exact values.
One can simply set the Interest limit based on the average values of round trip time and observed Data packet size, as the network buffers can smooth out most of the network fluctuations.

In the rest of the paper we will refer to Interest limits set based on the bandwidth-delay product as \emph{Physical limits}.
Pseudocode~\ref{alg:simple limits} shows implementation details of how the Physical limits algorithms.
Every time an Interest is ready to be send to an outgoing interface $of$ (OutInterest function), we check if the token bag is not empty.
If there is at least one free token available, we are ``borrowing'' it, forwarding the Interests, and remembering the outgoing interface in the corresponding entry of PIT.
Functions InData and Timeout ensure that tokens are returned whenever the requested Data is received or Interest times out.

\floatname{algorithm}{Pseudocode}

\begin{algorithm}[h]
\caption{Physical limits}
\label{alg:simple limits}
\begin{algorithmic}[1]
\For{\textbf{each} interface \em{f}}
    \State $L_{f} \leftarrow$ Interest Limit according to (1)
    \State $O_{f} \leftarrow 0$ \Comment{Outstanding Interests on interface \textbf{f}}
\EndFor

\vspace{0.2cm}
\Function{OutInterest}{Interest \textbf{i}, OutInterface \textbf{of}}
    \If{$L_{of} - O_{of} > 0$} \Comment{\textbf{of} is under physical limits}
        \State $O_{of} \leftarrow O_{of} + 1$  \Comment{``Borrow'' token}
        \State add \textbf{of} to PIT entry and forward \textbf{i} to \textbf{of}
    \Else
        \State drop Interest
    \EndIf
\EndFunction

\vspace{0.2cm}

\Function{InData}{Data \textbf{d}}
   \State lookup PIT entry \textbf{p} for data \textbf{d}
   \For{\textbf{each} outgoing interface \textbf{of} in \textbf{p}}
        \State $O_{of} \leftarrow O_{of} - 1$ \Comment{``Return'' token}
   \EndFor
\EndFunction

\vspace{0.2cm}

\Function{Timeout}{PIT entry $i$}
   \For{\textbf{each} outgoing interface \textbf{of} in \textbf{p}}
        \State $O_{of} \leftarrow O_{of} - 1$ \Comment{``Return'' token}
   \EndFor
\EndFunction

\end{algorithmic}
\end{algorithm}

% In our experiments Interest/Data size ratio is about 1:30
Considering the fact that normally the size of the Interests packets is significantly smaller than the size of the requested Data packets, a router needs to consume only a small portion of the upstream link (from A to B) to fully utilize the downstream link (from B to A).
The direct conclusion from this observation, confirmed by our evaluations in the Section~\ref{sec:evaluation}, is that if a router forwards only the right amount of Interests, both mechanisms of the Interests flooding attack will stop working: the forwarded Interests would neither congest the upstream nor exhaust router's memory resources.
At the same time, the Physical limits may become a new mechanisms to carry out a denial of service attacks.
For example, if the router A has forwarded 125 junk Interests, it will not be able to accept any other Interests neither from a legitimate user or an attacker, until these junk Interests start to expire (e.g., within one second).
This essentially reduces the probability of legitimate Interests getting through and can result in an effective low-rate denial of service attack (e.g., see Fig.~\ref{fig:small-scale-topo 3}).
